Should geophysicists use gravity anomaly or gravity disturbance?

### 2018-05-10 manuscript/letter.tex
review conclusion

We debate the conceptual differences between the gravity
disturbance and the gravity anomaly.
Our reasoning suggests that the gravity disturbance is the
more appropriated quantity for representing the gravity
effect produced by the gravity sources.
In summary, we point out that:

\begin{enumerate}
%\renewcommand{\theenumi}{(\arabic{enumi})}
\renewcommand{\theenumii}{\alph{enumii}}

\item Almost all interpretation techniques assume, implicitly or
directly, that the gravity data is harmonic (e.g., 
upward/downward continuation, 
data processing with equivalent layer,
conversions between gravity and magnetic data,
computation of vertical derivatives via Fourier and Hilbert
transforms). As a consequence, they implicitly or directly assume that
the gravity data approximate the gravity disturbance.

\item Almost all forward modelling techniques compute
the vertical component of the gravitational attraction 
exerted by the gravity sources at the observation points.
Notice that the gravity anomaly requires the computation of gravity
on the Geoid, which is generally within the topographic masses.
Hence, almost all these techniques in fact compute the gravity 
disturbance.

\item The gravity anomaly is defined as the difference between 
the gravity on the Geoid and the normal gravity on the reference 
ellipsoid, both at the same geodetic latitude and longitude.
Consequently, the gravity anomaly is a function of
the geodetic latitude and longitude only and cannot
be calculated at arbitrary heights. On the other hand,
the gravity disturbance can be computed at arbitrary points
outside the sources. Then, computing the upward 
continuation of gravity anomalies implicitly assumes 
that they approximate the gravity disturbance.

\item In Brazil, the maximum difference between the gravity 
disturbance (on or close to the Earth's surface) and the Free-air
anomaly reaches $\approx 9.3$ mGal.

\end{enumerate}

The principal theoretical implication of this study is that,
although the gravity anomaly may be used as a good approximation of
the gravitational effect produced by the sources for most practical applications,
the more appropriated quantity for gravity modelling is the
gravity disturbance. We stress that, more important than the 
terminology changes, the researchers must bear in mind the 
conceptual assumptions used in gravity modelling.

@@

\end{document}

>>>>

Although the gravity disturbance is a well-known quantity in geodesy,
its use in the applied geophysics is uncommon. 
For years, geophysicists have been using gravity anomaly for modelling
the gravitational effects due to geological bodies.
The question "Should geophysicists use gravity anomaly or gravity
disturbance for modelling the gravitational effect of gravity sources?"
is of more than academic interest.

Theoretically, the gravity anomaly is the difference between the
gravity on the Geoid and the normal gravity on the reference ellipsoid,
both at the same geodetic latitude and longitude.
On the other hand, the gravity disturbance is defined by the difference
between gravity and normal gravity at the same point.
Because these quantities are at the same point, the gravity disturbance
is a purely gravitational effect due to the gravity sources whereas the
gravity anomaly is a combination of gravitational and centrifugal
effects.
Consequently, the gravity disturbance is a harmonic function whereas the
gravity anomaly is not.
Because the gravity disturbance is a purely gravitational effect, it can
be represented by a harmonic function and satisfies the premise required
by the most techniques for processing potential-field data (e.g.,
upward/downward continuation, data processing with equivalent layer,
conversions between gravity and magnetic data, computation of vertical
derivatives via Fourier and Hilbert transforms).
Hence, the gravity disturbance is the more appropriated quantity for
representing the gravity effect produced by the gravity sources.

In practical problems of applied geophysics, geophysicists perform the
gravity-forward modelling by computing the vertical component of the
gravitational attraction exerted by the sources at the observation
points, regardless of the Geoid or reference ellipsoid positions,
for example.
This means that, actually, they compute gravity disturbance to fit 
a gravity anomaly (e.g., Bouguer anomaly) and this is a clear inconsistency.
Moreover, upward or downward continuations, for example, are commonly applied to
gravity anomalies. In this case, the gravity anomaly is wrongly considered a
harmonic function and this is another inconsistency.

By adopting these practical procedures, geophysicists are implicitly
taking the gravity anomaly as a good approximation of the gravity
disturbance. But why should geophysicists care about this usual
procedure? Because it may result in wrong interpretations of
quantitative nature. In Brazil, for example, the maximum difference
between the gravity disturbance and the free-air anomaly reaches 9.3
mGal. Such difference on the Earth's surface may lead, for example, to
an error of about xxx m in the depth-to-center estimate of a sphere-like
source with true depth-to-center at xxx m, radius xxxx m and density
contrast of xxx $kg \, m^{-3}$.

@@

\end{document}\grid
